Road Segment Prioritization for Bicycle Infrastructure

### 2020-12-29 README.Rmd
conclusion edits

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment. The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city. Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

<!-- There are a couple of files that cannot be synced to github due to their -->
<!-- size. These files are neseccary for the scripts to run. Below are links -->
<!-- to where you can download them, and instructions on where to place them -->
<!-- in the repo file structure -->

<!-- Flow Data (2011 Census Origin-Destination Data): -->

<!--   - Source: <https://www.nomisweb.co.uk/census/2011/bulk/rOD1> —\> -->
<!--     Choose File **“WU03EW”** -->
<!--   - Location in Repo: data-raw/flow\_data.csv -->

@@

> "Inadequate cycling infrastructure discourages cycling and wastes public money. Much cycling infrastructure in this country is inadequate. It reflects a belief, con- scious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"

@@

A considerable amount of research has been done towards that end. Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators]. Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal]. However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far]. Another defining feature for cycling infrastructure is how well connected it is. Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining]. While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].

@@

It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study. It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs. Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city. More accurate routing could be carried out given the availability of road-level data. In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.

@@

1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it.

@@

table <- table %>% sf::st_drop_geometry() %>% 
  dplyr::group_by(Community) %>%
  summarize(`Person-Km (Total)` = round(sum(person_km),0),
            `Person-Km (%)` = round(sum(person_km_perc),1))

@@

<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->

The research presented in this paper provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders. The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census). It should be noted that the recommendations are only as good as the granularity of the available data.

@@

<!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
<!--  HUSSEIN TO DO!-->

<!-- There should also be links to the literature. -->

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment. The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city. Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

<!-- The following file cannot be synced to github due to its size. -->
<!-- The file is neseccary for the scripts to run. Below is a link -->
<!-- to where you can download it, and instructions on where to place it -->
<!-- in the repo file structure -->

@@

> "Inadequate cycling infrastructure discourages cycling and wastes public money. Much cycling infrastructure in this country is inadequate. It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"

@@

A considerable amount of research has been done towards that end. Segregated cycling infrastructure^[*Segregated cycling infrastructure* refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators]. Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal]. However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far]. Another defining feature for cycling infrastructure is how well connected it is. Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining]. While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].

@@

It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study. It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs. Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city. More accurate routing could be carried out given the availability of road-level data. In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.

@@

<!-- 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it. -->
```{r, eval = FALSE}
1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it.
```

@@

table <- table %>% sf::st_drop_geometry() %>% 
  dplyr::group_by(Community) %>%
  summarize(`Person-Km (Total)` = format(round(sum(person_km),0), big.mark=","),
            `Person-Km (%)` = round(sum(person_km_perc),1))

@@

<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->
<!-- (HM) Answer: done --> 
This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study.

The methodology is designed so as to respect both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders. It can can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]. It should be noted that the recommendations are only as good as the granularity of the available data.

@@

<!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
<!-- (HM) Answer: Done. I added a footnote to the first paragraph in the conclusion linking to the github repo -->

<!-- There should also be links to the literature. -->
<!-- (HM) Answer: There are links to the literature at the end of each thematic section (Calculating Potential Cycling Demand / Routing / Road Segment Prioritization). There is a discussion paragraph in each of them. What additional links are necessary here? -->
